---
title: "Pulling Datasets & Cohort Selection"
author: "Julia Shoobe"
date: '`r Sys.Date()`'
output: html_document
---

# This file pulls the necessary datasets from the CRDM. It also filters for the cohort of interest -
  1. ED admit date on/after July 1, 2024
  2. Hospital Abbreviation = DUH
  3. Had Lab3234
# It pulls the following datasets:
  1. Encounters
  2. Labs
  3. Demographics
  4. Diagnosis
  5. Procedures
  6. Medication Administration
  7. Vitals
  8. Unstructured Notes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r}
library(RODBC)
library(stringr)
library(tidyverse)
library(lubridate)
library(purrr)
library(progressr)
```

# Connect to DukeHealth Server
```{r}
conn <- odbcDriverConnect('driver={ODBC Driver 17 for SQL Server};server=mccsdb-prd.dhe.duke.edu;database=ACE_DATA;trusted_connection=yes')
```

# Cohort Selection
## Encounters Table
```{r}
# ED Admit Date on/after July 1, 2024
# Hospital Abbreviation = DUH
PCOR_RSH_ENC_IP_ED_VW <- sqlQuery(conn,
  "SELECT *
   FROM CRDM_Views.PCOR_RSH_ENC_IP_ED_VW AS DEM
   WHERE DEM.HOSPITAL_ABBR = 'DUH'
     AND DEM.ED_ADM_DATE >= '2024-07-01';"
)
# 82071 observations

write.csv(PCOR_RSH_ENC_IP_ED_VW, "Datasets/encounters.csv", row.names = FALSE)

# All the encounter IDs for encounters on/after July 1, 2024 at DUH
encounter_id <- PCOR_RSH_ENC_IP_ED_VW |> 
  select(ENCOUNTERID)

# All the unique patient IDs for patients who visited DUH on/after July 1, 2024
patient_id <- PCOR_RSH_ENC_IP_ED_VW |> 
  distinct(PATID)
         # 56808 unique PATIDs
```

# Labs Table
```{r}
# Only include LAB3234, LAB3235, LAB3236, LAB3237 
PCOR_RSH_LAB_VW <- sqlQuery(conn,
  "SELECT *
   FROM CRDM_Views.PCOR_RSH_LAB_VW DEM
   WHERE DEM.LAB_PX IN ('LAB3234', 'LAB3235', 'LAB3236', 'LAB3237');"
) 

write.csv(PCOR_RSH_LAB_VW, "Datasets/labs.csv", row.names = FALSE)

# The most recent lab date
max(PCOR_RSH_LAB_VW$LAB_ORDER_DATE) # 2025-09-29

# Only includes encounters that have LAB3234 and the encounters must be at DUH on/after July 1, 2024
cohort <- encounter_id |> 
  inner_join(PCOR_RSH_LAB_VW, by = "ENCOUNTERID") |> 
  select(ENCOUNTERID, PATID, RESULT_NUM, RAW_LAB_NAME, RAW_RESULT)
# 38725 observations

write.csv(cohort, "cohort.csv", row.names = FALSE)

# The unique PatientIDs from patients who had the troponin lab, 
unique_patid_troponin <- cohort |> 
  distinct(PATID) # 11136

# unique ENCOUNTERIDs for troponin tests
unique_encounterid_trop <- cohort |> 
  distinct(ENCOUNTERID) # 13648
```

## Demographics Table
```{r}
PCOR_RSH_DEMOGRAPHIC_VW <- sqlQuery(conn,
  "SELECT
      DEM.PATID,
      DEM.SEX,
      DEM.RAW_RACE,
      DEM.BIRTH_DATE,
      DEM.HISPANIC,
      DEM.RAW_PAT_PREF_LANGUAGE_SPOKEN
   FROM CRDM_Views.PCOR_RSH_DEMOGRAPHIC_VW DEM;")


# demographics for unique PATIDs
dem_unique <- PCOR_RSH_DEMOGRAPHIC_VW |> 
  semi_join(patient_id, by = "PATID")

write.csv(dem_unique, "Datasets/demographics.csv", row.names = FALSE)
```

## Diagnostics Table
```{r}
PCOR_RSH_DIAGNOSIS_VW <- sqlQuery(conn,
  "SELECT
      DEM.ENCOUNTERID,
      DEM.ENC_TYPE,
      DEM.DX
    FROM CRDM_Views.PCOR_RSH_DIAGNOSIS_VW DEM
    WHERE DEM.ENC_TYPE = 'ED';"
  )

write.csv(PCOR_RSH_DIAGNOSIS_VW, "Datasets/diagnosis.csv", row.names = FALSE)
```

## Conditions Table
```{r}
PCOR_RSH_CONDITION_VW <- sqlQuery(conn,
  "SELECT
      DEM.PATID,
      DEM.ENCOUNTERID,
      DEM.CONDITION,
      DEM.REPORT_DATE,
      DEM.RESOLVE_DATE,
      DEM.ONSET_DATE
    FROM CRDM_Views.PCOR_RSH_CONDITION_VW DEM
  ")

write.csv(PCOR_RSH_CONDITION_VW, "Datasets/conditions.csv", row.names = FALSE)
```

```{r}
patids <- unique(unique_patid_troponin$PATID)
chunk_size <- 1000

# Batch IDs
patid_chunks <- split(patids, 
                      ceiling(seq_along(patids) / chunk_size)
                      )
```

# Procedures Table
```{r}
handlers("txtprogressbar")

PCOR_RSH_PROCEDURE_VW <- with_progress({
  p <- progressor(along = patid_chunks)
  
  map_dfr(patid_chunks, function(chunk) {
    p()
  
  chunk <- gsub("'", "''", chunk)
  chunk_sql <- paste0("'", chunk, "'", collapse = ",")
  
  sqlQuery(
    conn,
    paste0(
      "
      SELECT
          DEM.PATID,
          DEM.ENCOUNTERID,
          DEM.PX
      FROM CRDM_Views.PCOR_RSH_PROCEDURE_VW DEM
      WHERE DEM.PATID IN (", chunk_sql, ");
      "
    )
  )
  })
})

write.csv(PCOR_RSH_PROCEDURE_VW, "Datasets/procedures.csv", row.names = FALSE)
```

# Medication Administration Table
```{r}
ids <- cohort$ENCOUNTERID

id_list <- paste(sprintf("'%s'", ids), collapse = ", ")

query <- sprintf("
    SELECT
      ENCOUNTERID,
      PATID,
      MEDADMIN_ROUTE,
      RAW_MEDADMIN_MED_NAME,
      MEDADMIN_START_DATE,
      MEDADMIN_START_TIME
   FROM CRDM_Views.PCOR_RSH_MED_ADMIN_VW
   WHERE ENCOUNTERID IN (%s);
  ", id_list)

med_admin <- sqlQuery(conn, query)

write.csv(med_admin, "Datasets/med_admin.csv", row.names = FALSE)
```

# Vitals Table
```{r}
PCOR_RSH_VITAL_VW <- with_progress({
  p <- progressor(along = patid_chunks)
  
  map_dfr(patid_chunks, function(chunk) {
    p()
  
    # Build SQL IN clause for chunk
    chunk <- gsub("'", "''", chunk)
    chunk_sql <- paste0("'", chunk, "'", collapse = ",")
  
    # Query database
    sqlQuery(
      conn,
      paste0(
        "
        SELECT
          ENCOUNTERID, 
          PATID,
          SYSTOLIC,
          DIASTOLIC
          FROM CRDM_Views.PCOR_RSH_VITAL_VW
      WHERE PATID IN (", chunk_sql, ");
        "
      )
    )
  })
})

write.csv(PCOR_RSH_VITAL_VW, "Datasets/vitals.csv", row.names = FALSE)
```

## Notes Table
```{r}
encounterids <- unique(unique_encounterid_trop$ENCOUNTERID)

# Batch IDs
encounterid_chunks <- split(encounterids, ceiling(seq_along(encounterids) / 100))

notes <- with_progress({
  p <- progressor(along = encounterid_chunks)
  
  map_dfr(encounterid_chunks, function(chunk) {
    p()
    
    chunk <- gsub("'", "''", chunk)
    chunk_sql <- paste0("'", chunk, "'", collapse = ",")
    
    query <- paste0(
      "SELECT * FROM CRDM_Views.PCOR_RSH_NOTES_VW ",
      "WHERE ENCOUNTERID IN (", chunk_sql, ");"
    )
    
    dat <- sqlQuery(conn, query, stringsAsFactors = FALSE)
    
    dat
  })
})

unique(notes$REPORT_TYPE)

matches <- grepl("ED", as.character(notes$REPORT_TYPE), fixed = TRUE)
notes[matches, ]


write.csv(notes, "notes.csv", row.names = FALSE)
notes_data <- read.csv("Datasets/notes.csv", stringsAsFactors = FALSE)
# query_preview <- "SELECT TOP 100 * FROM CRDM_Views.PCOR_RSH_NOTES_VW;"
# notes_preview <- sqlQuery(conn, query_preview, stringsAsFactors = FALSE)

ed_provider_notes <- notes_data |> 
  filter(REPORT_TYPE == "ED Provider Notes")

write.csv(ed_provider_notes, "Datasets/ed_provider_notes.csv", row.names = FALSE)
```

# MRN Table
```{r}
ids <- cohort$PATID

patid_list <- paste(sprintf("'%s'", ids), collapse = ", ")

query <- sprintf("
    SELECT DEM.PATID, DEM.MRN
   FROM CRDM_Views.PCOR_RSH_MRN_XREF_VW AS DEM
   WHERE DEM.PATID IN (%s);
  ", patid_list)

PCOR_RSH_MRN_XREF_VW <- sqlQuery(conn, query)

write.csv(PCOR_RSH_MRN_XREF_VW, "Datasets/mrns.csv", row.names = FALSE)
```
